---
title: "Data Analysis"
author: "Caroline & Eli"
date: "4/16/25"
format: html
execute: 
  warning: false
  message: false
---

# Data Analysis

## Introduction and Data
	In the United States, climate change has become a heavily contentious and debated issue. The climate crisis is one of the largest issues that we are facing today, and understanding the general population's stance and views on climate change is important for policy making decisions. There have been a myriad of studies looking at what factors are correlated with a belief in climate change, like Robinson and Furnham’s (2022) study: Correlates of belief in climate change: Demographics, ideology and belief systems. There are also studies investigating companies' willingness to pay to reduce climate change, like Zhao et. all (2018): Factors influencing companies' willingness to pay for carbon emissions: Emission trading schemes in China. However, there have not been very many studies examining what impacts individuals willingness to pay to counteract climate change. This is a very important area to examine since public support for climate change action is necessary to influence governments to make changes - like public support for carbon reduction policies which we will examine later. 
	The question we aim to answer throughout this data analysis is: How do individual beliefs (regarding climate change, political stances, education levels, and views on the role of government) influence personal willingness to spend to reduce climate change?
	In this dataset we anticipate that there will be a strong positive correlation between individuals that are more liberal and higher educated and willingness to pay to reduce climate change. We also anticipate the reverse, that individuals that are less educated and more conservative will have lower willingness to pay to support climate change. We also expect that the more an individual believes in the reality of global warming and the more pressing they find the issue to be, the more likely they are to be willing to pay to reduce greenhouse gas emissions. 

The data set that we are analyzing comes from The National Surveys on Energy and Environment (NSEE). The NSEE is a biannual national opinion survey. This data set contains 18 survey batches collected from Fall 2008 to Spring 2017. Each case represents an individual’s responses to the survey questions. There are 15,136 total cases. 

## Relevant variables:
CARBONTAX - These are 3 variables that look at different levels of willingness to pay in support of the Carbon tax in this general format: Support for Carbon fuels tax if it raises costs by X$ a month. What if the carbon fuels taxes significantly lowered greenhouse gases but increased your energy costs by X dollars a month? Would you strongly support, somewhat support, somewhat oppose or strongly oppose this type of system?
STATE_RPS_ECONOMY- This measures the level of agreement respondents have with this statement- State governments will boost their economies by requiring greater use of renewable energy.
INTL_ECONOMICIMPACT- This measures the level of agreement respondents have with this statement- Overall, the U.S. economy would benefit from leaving the international climate treaty designed to reduce greenhouse gas emissions)
DEMOG_POLVIEWS: This reports respondents Political beliefs, on a scale from conservative to liberal
GW_CONCERN: This reports how concerned respondents are about the issue of global warming.
BELIEVER_IMMEDIACY: This reports if respondents believe that global warming requires immediate government action.
GOVT_FED: This measures the level of agreement respondents have with this statement- How much responsibility does the federal government have for taking actions to reduce global warming?
DEMOG_EDU: The Highest level of education respondents have. 

## Data wrangling
The majority of our data cleaning entailed removing NA values (there were a lot since not all questions were asked each year), as well as Not Sure or Refused to answer (Since they don’t provide any real information into how the respondent stands on economic or climate based issues).

 In terms of data wrangling - the majority of it came prior to graphing and the creation of models - many variables had string responses which made it harder to read comparisons. We had to convert the variable INTL_ECONOMICIMPACT from String responses to numerical responses in order to take an average value of the responses for the different BELIEVER_IMMEDIACY values. We also simplified WAVE_NUMBER into just the year that the surveys were taken and converted them into `Date` data types for ease of graphing. January 1st was given as an arbitrary date and does not represent the true date that the surveys were taken. For our model, we simplified the variable DEMOG_POLVIEWS to only have 3 political standings: We combined Very liberal and Somewhat liberal to just be Liberal, and combined Very Conservative and Somewhat Conservative to just be Conservative, and left Moderate the same. This was necessary in order to allow the decision tree to be readable - and to have enough data at each end node. 

We created one important variable that we will use in our analysis: 
CARBONTAX_SUPPORT: This variable takes in the 3 levels of willingness to pay for the carbon tax, and returns the maximum value a given individual will be willing to pay for the carbon tax - or if they are unwilling to pay at all. 

In our model, we plan to use CARBONTAX_SUPPORT as our outcome variable that will be predicted by our decision tree. This was the variable in our data set that we thought was the best for showing an individual’s willingness to pay to reduce climate change, since this tax explicitly attempts to reduce greenhouse gas emissions by increasing the prices of fossil fuels. So, it is our best measure of what individuals care about more, the environment, or their own financial standing. We thought that two of the more important factors that will determine willingness to pay for carbon reduction practices would be if they do or do not believe in global warming in the first place - and their political ideologies, since global warming has increasingly become a political issue. We are using GW_BELIEF (if individuals believe in global warming or not), and DEMOG_POLVIEWS (the respondents political beliefs), to predict what level of the Carbon tax individuals will be willing to pay - we model this using a decision tree. 

## Methodology 
  
```{r install packages}
#| output: true

#Install Packages
library(polite)
library(rvest)
library(tibble)
library(tidyverse)
library(dplyr)
```

```{python}
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns
from sklearn.preprocessing import StandardScaler, OneHotEncoder
from sklearn.pipeline import Pipeline
from sklearn.compose import make_column_transformer
from sklearn.linear_model import LinearRegression 
from sklearn.model_selection import cross_val_predict, train_test_split, cross_val_score 
from sklearn.metrics import mean_squared_error, r2_score
from sklearn.linear_model import Ridge, RidgeCV, Lasso, LassoCV, ElasticNetCV
from sklearn.tree import DecisionTreeClassifier, DecisionTreeRegressor, plot_tree
from sklearn.ensemble import RandomForestClassifier, RandomForestRegressor
from sklearn.ensemble import GradientBoostingClassifier, GradientBoostingRegressor
```


```{r load data}
#Loading the dataset
load("36368-0001-Data.rda")
econ_clim_data <- as.data.frame(da36368.0001)
```

```{r new dataframe}
#Creating a dataframe with only our necessary variables

working_dataset <- econ_clim_data %>%
  select("WAVE_NUMBER", "INTL_ECONOMICIMPACT", "STATE_VEHICLES_V1",
         "STATE_VEHICLES_V2", "CARBONTAX_COST10PCT", "CARBONTAX_COST15DOL",
         "CARBONTAX_COST50DOL",  "STATE_RPS_ECONOMY", "GW_BELIEF",
         "GW_CONCERN", "BELIEVER_IMMEDIACY", "GOVT_EFFECTIVENESS", "GOVT_FED",
         "STATE_NEIGHBORSACT", "GASTAX_GOVT", "FFTAX_GOVT",
         "CAPTRADE_GOVT_V1", "GOVT_EFFECTIVENESS_V2", "GOVT_EFFECTIVENESS",
         "STATE_RPS_V1", "STATE_RPS_V2", "STATE_RPS_V2_COST100",
         "STATE_RPS_V3", "STATE_RPS_V3_COST25", "STATE_RPS_V3_COST50",
         "DEMOG_EDU", "DEMOG_POLVIEWS", "DEMOG_RELIG")
```

```{r summary table}
#Summary Table for our dataset
summary_table <- summary(working_dataset)
summary_table
```

```{r demog_polviews_state_rps_economy_graph}
# count instances unique values in STATE_ECO and
# save it to a new data structure to graph easier
eco_by_polview <- as.data.frame(table(working_dataset$DEMOG_POLVIEWS,
                                      working_dataset$STATE_RPS_ECONOMY))
colnames(eco_by_polview) <- c("POLVIEWS", "STATE_ECO", "Count")

eco_by_polview <- eco_by_polview %>%
  spread(key = STATE_ECO, value = Count, fill = 0)

eco_by_polview <- eco_by_polview %>%
  pivot_longer(cols = c("(1) Strongly agree", "(2) Somewhat agree",
                        "(3) Somewhat disagree", "(4) Strongly disagree"),
               names_to = "STATE_ECO",
               values_to = "Count")

polviews_state_eco <- eco_by_polview %>%
  ggplot(aes(x=POLVIEWS,
             y=Count,
             fill = STATE_ECO)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(x="Political Views",
       y="Count of Respondents",
       title = "Suport for States Investing in Renewable Energies
                to Benefit their Economies") +
  guides(fill=guide_legend(title="Do renewable energies
grow economies?"))
```

```{r poliews_state_eco_show}
#| echo: false
polviews_state_eco
```

This graph shows the level of support respondents of varying political ideologies have for states investing in renewable energies based on perceived economic results.

```{r clean dates}
# remove everything but year info from WAVE_NUMBER
working_dataset <- working_dataset %>%
  mutate(WAVE_NUMBER = (sub('.*(?=.....$)', '', WAVE_NUMBER, perl=T)) %>%
                        paste("/01/01", sep="")
                        )
```

```{r date and international economic impact}
# new dataframe with only observations
# that answered the STATE_RPS_ECONOMY question
date_intl_graph <- working_dataset[working_dataset$STATE_RPS_ECONOMY %in%
                                   c("(1) Strongly agree",
                                     "(2) Somewhat agree",
                                     "(3) Somewhat disagree",
                                     "(4) Strongly disagree"),]

# stacked bar graph
date_intl_eco <- date_intl_graph %>%
  ggplot(aes(x=WAVE_NUMBER,
             y=STATE_RPS_ECONOMY,
             fill = STATE_RPS_ECONOMY)) + 
  geom_col(position = "stack") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        axis.text.y = element_blank()) +
  labs(x = "Year",
       y = "Count of Respondents",
       title = "Suport for States Investing in Renewable Energies
       to Benefit their Economies Over Time") +
  guides(fill=guide_legend(title="Do renewable energies
grow economies?"))
```
```{r date_intl_eco_show}
#| echo: false
date_intl_eco
```

This graph shows the change in respondent's beliefs about if states investing in renewable energies would be beneficial to their economies or not.

```{r carbontax-concern-graph}
#Graph 1: Support for Carbon Tax by Levels of Concern About Global Warming
#Getting rid of NA values in GW_Concern
excluded_values <- c(NA)

# Calculate percentages of people that strong support, somewhat support,
# somewhat oppose, and strongly oppose for differnet GW_Concern responses
# while removing NA and "Not Sure" responses
data_carbontax10 <- working_dataset %>%
  filter(!GW_CONCERN %in% excluded_values, !is.na(CARBONTAX_COST10PCT)) %>% 
  group_by(GW_CONCERN, CARBONTAX_COST10PCT) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(GW_CONCERN) %>%
  mutate(Percentage = (Count / sum(Count)) * 100) %>%
  mutate(CarbonTax = "Revenue Neutral")

data_carbontax15 <- working_dataset %>%
  filter(!GW_CONCERN %in% excluded_values, !is.na(CARBONTAX_COST15DOL)) %>%
  group_by(GW_CONCERN, CARBONTAX_COST15DOL) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(GW_CONCERN) %>%
  mutate(Percentage = (Count / sum(Count)) * 100) %>%
  mutate(CarbonTax = "$15/month")

data_carbontax50 <- working_dataset %>%
  filter(!GW_CONCERN %in% excluded_values, !is.na(CARBONTAX_COST50DOL)) %>%  
  group_by(GW_CONCERN, CARBONTAX_COST50DOL) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(GW_CONCERN) %>%
  mutate(Percentage = (Count / sum(Count)) * 100) %>%
  mutate(CarbonTax = "$50/month")

# Rename response column
colnames(data_carbontax10)[2] <- "Response"
colnames(data_carbontax15)[2] <- "Response"
colnames(data_carbontax50)[2] <- "Response"

# Combine data to use for the graph
data_combined <- bind_rows(data_carbontax10,
                           data_carbontax15,
                           data_carbontax50)

# Creating the stacked bar chart
ggplot(data_combined, aes(x = GW_CONCERN, y = Percentage, fill = Response)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ CarbonTax, ncol = 1) +  
  labs(
    title = "Support for Carbon Tax by Levels of Concern About Global Warming",
    x = "Level of Concern About Global Warming",
    y = "Percentage of Respondants",
    fill = "Response"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 35, hjust = 1, size = 5))
```

This graph shows the level of support respondents have for the different variations of the Carbon tax which are designed to significantly lower greenhouse gases. Revenue Neutral means that every dollar collected by the government would be returned to the public as an income tax rebate. 15\$/month means it would increase energy costs by 15 dollars a month, and 50\$/month means it would increase energy costs by 50\$/month. These charts show that the more concerned an individual is about global warming, the more they will support the various levels of carbon tax - and the support decreases as costs go up. The largest support for carbon tax is found in individuals who are very concerned about global warming at the revenue neutral level - and there is almost no support found for individuals who are not concerned about global warming at all at any level of cost.  

```{r econ-impact-climate-belief}
#Graph 2: Economic Impact Perception vs. Belief in Climate Immediacy
#Removing NA values for graph
clean_data <- working_dataset %>%
  filter(!is.na(INTL_ECONOMICIMPACT), !is.na(BELIEVER_IMMEDIACY))

#Changing INTL_ECONOMICIMPACT to have numerical values instead of words
second_graph <- clean_data %>%
  filter(INTL_ECONOMICIMPACT %in% c(
    "(1) Strongly agree", "(2) Somewhat agree", 
    "(3) Somewhat disagree", "(4) Strongly disagree"
  )) %>%
  mutate(INTL_ECONOMICIMPACT = case_when(
    INTL_ECONOMICIMPACT == "(1) Strongly agree" ~ 1,
    INTL_ECONOMICIMPACT == "(2) Somewhat agree" ~ 2,
    INTL_ECONOMICIMPACT == "(3) Somewhat disagree" ~ 3,
    INTL_ECONOMICIMPACT == "(4) Strongly disagree" ~ 4
  ))

#Calculating the average numerical value of INTL_ECONOMICIMPACT
# for each response of BELIEVER_IMMEDIACY
result <- second_graph %>%
  group_by(BELIEVER_IMMEDIACY) %>%
  summarise(avg_intl_economic_impact = mean(INTL_ECONOMICIMPACT, na.rm = TRUE))

# Plotting the result
ggplot(result, aes(x = BELIEVER_IMMEDIACY,
                   y = avg_intl_economic_impact,
                   group = 1)) +
  geom_line(color = "#1f78b4", size = 1.5) +   
  geom_point(linewidth = 2.5, color = "purple") +   
  labs(
    title = "Economic Impact Perception vs. Belief in Climate Immediacy",
    x = "Climate Change Requires Immediate Government Action",
    y = "Leaving Climate Treaty Benefits Economy"
  ) +
  scale_y_continuous(limits = c(1, 4),
    breaks = seq(1, 4, by = 1),
    labels = c("Strongly Agree", "Somewhat Agree",
               "Somewhat Disagree", "Strongly Disagree")
  ) +
  theme_bw(base_size = 11.5)+
  theme(plot.margin = margin(10, 20, 10, 20))
```

This graph shows the average level that respondents believe that the U.S. economy would benefit from leaving the international climate treaty designed to reduce greenhouse gas emissions - if they believe or don't believe that global warming requires immediate government action. This graph shows that if an individual believes that climate change requires immediate government action, then they are more likely to believe that leaving the climate treaty will hurt the US economy - not help it.
```{r creating new variables for graphing and analysis}
working_dataset2 <- working_dataset %>%
  mutate(
    CARBONTAX_SUPPORT = case_when(
      !is.na(CARBONTAX_COST50DOL) & CARBONTAX_COST50DOL %in% c("(1) Strongly support", "(2) Somewhat support") ~ "Willing to pay $50/month",
      !is.na(CARBONTAX_COST15DOL) & CARBONTAX_COST15DOL %in% c("(1) Strongly support", "(2) Somewhat support") ~ "Willing to pay $15/month",
      !is.na(CARBONTAX_COST10PCT) & CARBONTAX_COST10PCT %in% c("(1) Strongly support", "(2) Somewhat support") ~ "Willing to pay Revenue Neutral",
      !is.na(CARBONTAX_COST50DOL) & CARBONTAX_COST50DOL %in% c("(3) Somewhat oppose", "(4) Strongly oppose") |
  	!is.na(CARBONTAX_COST15DOL) & CARBONTAX_COST15DOL %in% c("(3) Somewhat oppose", "(4) Strongly oppose") |
  	!is.na(CARBONTAX_COST10PCT) & CARBONTAX_COST10PCT %in% c("(3) Somewhat oppose", "(4) Strongly oppose") ~ "Not willing to pay"
    )
  )

working_dataset2 <- working_dataset2 %>%
  mutate(
    DEMOG_POLVIEWS = case_when(
      DEMOG_POLVIEWS %in% c("(4) Somewhat liberal", "(5) Very liberal") ~ "Liberal",
      DEMOG_POLVIEWS %in% c("(2) Somewhat conservative", "(1) Very conservative") ~ "Conservative",
      DEMOG_POLVIEWS %in% c("(3) Moderate") ~ "Moderate",
      TRUE ~ NA_character_
    )
  )

# View the dataset with the new variable
```

```{r another graph}
#Graph showing the percent support for Carbon Tax by the Level of Responsibility they think the Government should have over Climate change and their politcal Views
working_dataset2 %>%
  filter(!is.na(CARBONTAX_SUPPORT), !is.na(GOVT_FED), !is.na(DEMOG_POLVIEWS)) %>%
  mutate(SUPPORT_BINARY = ifelse(CARBONTAX_SUPPORT %in% c("Willing to pay $50/month", 
                                                           "Willing to pay $15/month", 
                                                           "Willing to pay Revenue Neutral"), 1, 0)) %>%
  group_by(GOVT_FED, DEMOG_POLVIEWS) %>%
  summarise(mean_support = mean(SUPPORT_BINARY, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = DEMOG_POLVIEWS, y = GOVT_FED, fill = mean_support)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "darkgreen", labels = scales::percent) +
  labs(
    x = "Political Views",
    y = "Federal Gov Level Responsibility",
    fill = "% Support for Carbon Tax",
    title = "Support for Carbon Tax by Gov Responsibility & Political Views"
  ) +
  theme_minimal()
```
This graph shows the percent of individuals that support the carbon tax by their political ideology, and how much responsibility they believe the Federal Government has for taking action to reduce global warming. This shows that across political ideologies- the more an individual believes it is the responsibility of the government to reduce global warming, the more likely they are to be willing to pay for the carbon tax. Additionally- overall the more liberal an individual is the more likely they are to be willing to pay for the carbon tax.

## Results


## Discussion


## Citations
Rabe, Barry, et al. “National Surveys on Energy and the Environment, [United States], Fall 2008-Fall 2017.” National Surveys on Energy and the Environment, [United States], Fall 2008-Fall 2017, Inter-university Consortium for Political and Social Research [distributor], 24 Aug. 2020, www.icpsr.umich.edu/web/ICPSR/studies/36368. 

Teyton, Anaïs, and David M Abramson. “The Formation of Belief: An Examination of Factors That Influence Climate Change Belief among Hurricane Katrina Survivors.” Environmental Justice (Print), U.S. National Library of Medicine, 1 June 2021, pmc.ncbi.nlm.nih.gov/articles/PMC8259080/. 

Zhao, Yibing, et al. “Factors Influencing Companies’ Willingness to Pay for Carbon Emissions: Emission Trading Schemes in China.” Energy Economics, North-Holland, 5 Sept. 2018, www.sciencedirect.com/science/article/pii/S0140988318303694. 


