---
title: "Project Proposal"
Names: "Caroline & Eli"
format: html
editor: visual
---

## Data Set 1:

```{r}
#Install Packages
library(polite)
library(rvest)
library(tibble)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{r}
#Loading the dataset
load("/Users/siekmanc/Downloads/ICPSR_36368 2/DS0001/36368-0001-Data.rda")
econ_clim_data <- as.data.frame(da36368.0001)
View(econ_clim_data)
```

```{r}
#Creating a dataframe with only our necessary variables

working_dataset <- econ_clim_data %>% select("INTL_ECONOMICIMPACT", "STATE_VEHICLES_V1","STATE_VEHICLES_V2","CARBONTAX_COST10PCT","CARBONTAX_COST15DOL","CARBONTAX_COST50DOL", "STATE_RPS_ECONOMY","GW_BELIEF","GW_CONCERN","BELIEVER_IMMEDIACY","GOVT_EFFECTIVENESS","GOVT_FED","STATE_NEIGHBORSACT","GASTAX_GOVT","FFTAX_GOVT","CAPTRADE_GOVT_V1","GOVT_EFFECTIVENESS_V2","GOVT_EFFECTIVENESS","STATE_RPS_V1","STATE_RPS_V2","STATE_RPS_V2_COST100","STATE_RPS_V3","STATE_RPS_V3_COST25","STATE_RPS_V3_COST50","DEMOG_EDU","DEMOG_POLVIEWS","DEMOG_RELIG")

View(working_dataset)
```

```{r}
#Summary Table for our dataset
summary_table <- summary(working_dataset)
View(summary_table)
#Not sure if this is the absolute best looking method but it does provide a complete summary
```

```{r}
#Graph 1: Support for Carbon Tax by Levels of Concern About Global Warming
#Getting rid of NA values in GW_Concern
excluded_values <- c(NA)

# Calculate percentages of people that strong support, somewhat support, somewhat oppose, and strongly oppose for differnet GW_Concern responses while removing NA and "Not Sure" responses
data_carbontax10 <- working_dataset %>%
  filter(!GW_CONCERN %in% excluded_values, !is.na(CARBONTAX_COST10PCT)) %>% 
  group_by(GW_CONCERN, CARBONTAX_COST10PCT) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(GW_CONCERN) %>%
  mutate(Percentage = (Count / sum(Count)) * 100) %>%
  mutate(CarbonTax = "Revenue Neutral")

data_carbontax15 <- working_dataset %>%
  filter(!GW_CONCERN %in% excluded_values, !is.na(CARBONTAX_COST15DOL)) %>%
  group_by(GW_CONCERN, CARBONTAX_COST15DOL) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(GW_CONCERN) %>%
  mutate(Percentage = (Count / sum(Count)) * 100) %>%
  mutate(CarbonTax = "$15/month")

data_carbontax50 <- working_dataset %>%
  filter(!GW_CONCERN %in% excluded_values, !is.na(CARBONTAX_COST50DOL)) %>%  
  group_by(GW_CONCERN, CARBONTAX_COST50DOL) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(GW_CONCERN) %>%
  mutate(Percentage = (Count / sum(Count)) * 100) %>%
  mutate(CarbonTax = "$50/month")

# Rename response column
colnames(data_carbontax10)[2] <- "Response"
colnames(data_carbontax15)[2] <- "Response"
colnames(data_carbontax50)[2] <- "Response"

# Combine data to use for the graph
data_combined <- bind_rows(data_carbontax10, data_carbontax15, data_carbontax50)

# Creating the stacked bar chart
ggplot(data_combined, aes(x = GW_CONCERN, y = Percentage, fill = Response)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ CarbonTax, ncol = 1) +  
  labs(
    title = "Support for Carbon Tax by Levels of Concern About Global Warming",
    x = "Level of Concern About Global Warming",
    y = "Percentage of Respondants",
    fill = "Response"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This graph shows the level of support respondents have for the different variations of the Carbon tax which are designed to significantly lower greenhouse gases. Revenue Neutral means that every dollar collected by the government would be returned to the public as an income tax rebate. 15\$/month means it would increase energy costs by 15 dollars a month, and 50\$/month means it would increase energy costs by 50\$/month.

```{r}
#Graph 2: Economic Impact Perception vs. Belief in Climate Immediacy
#Removing NA values for graph
clean_data <- working_dataset %>%
  filter(!is.na(INTL_ECONOMICIMPACT), !is.na(BELIEVER_IMMEDIACY))

#Changing INTL_ECONOMICIMPACT to have numerical values instead of words
second_graph <- clean_data %>%
  filter(INTL_ECONOMICIMPACT %in% c(
    "(1) Strongly agree", "(2) Somewhat agree", 
    "(3) Somewhat disagree", "(4) Strongly disagree"
  )) %>%
  mutate(INTL_ECONOMICIMPACT = case_when(
    INTL_ECONOMICIMPACT == "(1) Strongly agree" ~ 1,
    INTL_ECONOMICIMPACT == "(2) Somewhat agree" ~ 2,
    INTL_ECONOMICIMPACT == "(3) Somewhat disagree" ~ 3,
    INTL_ECONOMICIMPACT == "(4) Strongly disagree" ~ 4
  ))

#Calculating the average numerical value of INTL_ECONOMICIMPACT for each response of BELIEVER_IMMEDIACY
result <- second_graph %>%
  group_by(BELIEVER_IMMEDIACY) %>%
  summarise(avg_intl_economic_impact = mean(INTL_ECONOMICIMPACT, na.rm = TRUE))

# Plotting the result
ggplot(result, aes(x = BELIEVER_IMMEDIACY, y = avg_intl_economic_impact, group = 1)) +
  geom_line(color = "#1f78b4", size = 1.5) +   
  geom_point(linewidth = 2.5, color = "purple") +   
  labs(
    title = "Economic Impact Perception vs. Belief in Climate Immediacy",
    x = "Climate Change Requires Immediate Government Action",
    y = "Leaving Climate Treaty Benefits Economy"
  ) +
  scale_y_continuous(limits = c(1, 4),
    breaks = seq(1, 4, by = 1),
    labels = c("Strongly Agree", "Somewhat Agree", "Somewhat Disagree", "Strongly Disagree")
  ) +
  theme_bw(base_size = 11.5)+
  theme(plot.margin = margin(10, 20, 10, 20))
```

This graph shows the average level that respondants believe that the U.S. economy would benefit from leaving the international climate treaty designed to reduce greenhouse gas emissions - if they believe or don't believe that global warming requires immediate government action.
